Parsing Postfix log files

### 2007-10-09 doc/logparser.tex
Added introduction and conclusion to other parsers section. Improved YAALA review.

@@

    \item [Yet Another Advanced Logfile Analyser] uses a plugin based
        system to analyse log files and produce output reports.  The core
        code is merely 91 lines long, as all the parsing and report
        generation is handled by modules.  Using YAALA as a base would be
        only slightly less work as both input and output modules would need
        to be written; it may even be more work to implement the parser
        within the constraints of YAALA\@.

        \url{http://yaala.org/} \newline (Last checked 2007/08/13.)

@@

>>>>

It is important to compare and contrast newly developed programs,
algorithms and parsers against those already available, to accurately judge
what, if any, improvements are delivered by the newcomers.  There are not
that many previously developed Postfix log parsers, indeed it was quite
difficult to find ten parsers to review for this project, and the
functionality offered ranges from quite basic to much more mature,
depending on the needs of the creator.  The programs are not reviewed from
an independent viewpoint; the objective is to compare and contrast each
program with \parsername{}.

@@

    \item [Yet Another Advanced Logfile Analyser] \textit{yaala is a very
        flexible analyser for all kinds of logfiles. It uses parsers to
        extract information from a logfile, an SQL-like query language to
        relate the information to each other and an output-module to format
        the information appropriately.\/}

        YAALA uses a plugin based system to analyse log files and produce
        HTML output reports.  The core code is merely 153 lines long, as
        all the parsing and report generation is handled by modules.  YAALA
        supports storing previously gathered data; it dumps its internal
        data structures using Perl's Storable module~\cite{perl-storable},
        so with some knowledge of the module responsible for parsing
        Postfix log files it should be possible to extract data with an
        external program.  Using YAALA as a base would be only slightly
        less work as both input and output modules would need to be
        written; it may even be more work to implement the parser within
        the constraints of YAALA\@.  YAALA bundles a Postfix parser, which
        uses hard-coded parsing to extract the following fields from
        specific log lines:

        Aggregations: count, bytes (the sum of bytes transferred; count is
        not explained).\newline
        Keyfields [sic]: date, hour, sender, recipient, defer\_count, delay,
        incoming\_host, outgoing\_host.

        YAALA provides an SQL like mini-language used when generating the
        reports; sample reports can be seen at~\cite{yaala-samples}.
        Example query: \newline \hspace*{2em} \texttt{requests BY file
        WHERE host =\~{} Google}. \newline The mini-language is quite
        limited, and cannot be used to extract data, merely to create
        reports.

        YAALA as distributed does not attempt the in-depth parsing
        explained in this document, and implementing it within the confines
        of YAALA would be at least as much work as implementing it
        separately.

        In summary YAALA provides a Postfix parser which is not extensible
        unless the user is prepared to change the source
        code,\footnote{\parsername{} requires source code changes for some
        new features, but parsing new log lines is accomplished by adding
        new rules.} and reasonably flexible report generation, but no
        facilities to extract data for use in other tools.

        \url{http://yaala.org/} \newline (Last checked 2007/10/09.)

@@

While there are other programs available which perform basic Postfix log
parsing (some to a greater level of intricacy and detail than others), none
attempt to correlate log lines by pid or queueid to produce an overall
record of the journey each mail traverses through Postfix.  Most other
parsers immediately generate a report and discard the data extracted from
the logs; those which don't discard the data retain it in a format
inaccessible to other tools.  All of the parsers reviewed can produce a
report of greater or lesser detail and complexity, a facility not offered
by \parsername{}; reporting is deferred to a subsequent application.

The overriding difference between \parsername{} and the other parsers
reviewed herein is that none of them aim to achieve the high level of
understanding of Postfix logs achieved by \parsername{}.
